---
title: "Basic Plot Types"
author: "Roy Thomas"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "Wilke-slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")


```

## Basic Plots using ggplot

---

### Load the needed libraries first
#### library(tidyverse)
#### library(palmerpenguins)
#### library(ggthemes)

```{r include = FALSE, echo = FALSE}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```


---

## Goal: Recreate this visualization

```{r penguin-final-graph, echo = FALSE, fig.width = 8.5, fig.asp = 0.5, dev = "svg"}
options(warn=-1)
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```

---
.small-font[
```{r}
glimpse(penguins)
```
]

---

### We need to tell ggplot how the data should be visually represented
### The mapping argument of the ggplot() function defines how variables in your dataset are mapped to visual properties (aesthetics) of your plot. 
### The mapping argument is always defined in the aes() function
### The x and y arguments of aes() specify which variables to map to the x and y 
---
### Putting all that together:
```{r fig.width = 7.5, fig.asp = 0.7}
ggplot(data = penguins, mapping = aes(x = flipper_length_mm, y = body_mass_g))

```

---
### This gave us an empty canvas
### We need to define the geometrical object, *geom*, that a plot uses to represent data.
### Bar charts use bar geoms (geom_bar()), line charts use line geoms (geom_line()), boxplots use boxplot geoms (geom_boxplot()), scatterplots use point geoms (geom_point()), and so on.
---
```{r fig.width = 7.5, fig.asp = 0.7}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point()

```

---

### When a categorical variable is mapped to an aesthetic, ggplot2 will automatically assign a unique value of the aesthetic (here a unique color) to each unique level of the variable (each of the three species), a process known as scaling. ggplot2 will also add a legend that explains which values correspond to which levels.
---

```{r fig.width = 7.5, fig.asp = 0.7}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g, color = species)) +
  geom_point()

```

---
### we will add a new geom as a layer on top of our point geom: geom_smooth(). And we will specify that we want to draw the line of best fit based on a linear model with method = "lm".
---

```{r fig.width = 7.5, fig.asp = 0.7}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")

```

---
#### When aesthetic mappings are defined in ggplot(), at the global level, they’re passed down to each of the subsequent geom layers of the plot. However, each geom function in ggplot2 can also take a mapping argument, which allows for aesthetic mappings at the local level that are added to those inherited from the global level. Since we want points to be colored based on species but don’t want the lines to be separated out for them, we should specify color = species for geom_point() only.
---
```{r fig.width = 7.5, fig.asp = 0.6}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species)) + 
  geom_smooth(method = "lm")

```

---
#### It’s generally not a good idea to represent information using only colors on a plot, as people perceive colors differently due to color blindness or other color vision differences. Therefore, in addition to color, we can also map species to the shape aesthetic.
---
```{r fig.width = 7.5, fig.asp = 0.6}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = species)) + 
  geom_smooth(method = "lm")

```



---
#### Now to add some labels

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = species)) + 
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()

```

---
```{r fig.width = 7.5, fig.asp = 0.5, echo=FALSE, include=TRUE}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = species)) + 
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()

```

---

#### To examine the distribution of a categorical variable, you can use a bar chart.


```{r fig.width = 7.5, fig.asp = 0.71}
penguins |> 
  ggplot(aes(x = species)) +
  geom_bar()

```

---

#### To reorder the bars based on their frequencies, first transform the variable to a factor (how R handles categorical data) and then reorder the levels of that factor.


```{r fig.width = 7.5, fig.asp = 0.5, echo=FALSE, include=TRUE}
penguins |> 
  ggplot(aes(x = fct_infreq(species))) +
  geom_bar()

```


---

## Numeric Variable
#### They are quantitative
#### They can take on a wide range of values
#### You can add/subtract/average etc
#### They can be discrete or continuous
#### One commonly used visualization for distributions of continuous variables is a histogram

---

## Histograms

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g)) +
  geom_histogram(binwidth = 200)

```

---

## Histograms - too narrow a binwidth

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g)) +
  geom_histogram(binwidth = 20)

```

---
## Histograms - too wide a binwidth

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g)) +
  geom_histogram(binwidth = 2000)

```

---
## Alternative to Histogram - Density Plot
#### A smoothed out version of the histogram

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g)) +
  geom_density()
```
---
## Watch out! Difference between "color" and "fill"

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
ggplot(aes(x = species)) +
  geom_bar(color = "red")

```
---

## Watch out! Difference between "color" and "fill"

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
ggplot(aes(x = species)) +
  geom_bar(fill = "red")

```

---
## Visualizing relationships
#### Between a numeric and categorical variable
#### Use side-by-side boxplots
---

## What is a box plot?
.center[
<img src = "basicPlotTypes_files/BoxPlotExplained.png", width = 50%, style = "margin-top: -0.5em; margin-bottom: -1em"></img>
]

---

#### Look at distribution of body mass by species, using boxplot


```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(species, body_mass_g)) +
  geom_boxplot()
```

---

#### Alternatively, doing a Density Plot

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g, color = species)) +
  geom_density()
```

---

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g, color = species)) +
  geom_density(linewidth = 1)
```

---

#### We can map color and fill to species
#### The alpha aesthetic adds transparency to the filled density curves
#### alpha=0 is completely transparent, alpha=1 is completely opaque


```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(x=body_mass_g,color = species, fill = species)) +
  geom_density(alpha = 0.5)

```

---

## Two Categorical Variables
#### Use stacked bar plots to visualize the relationship between two categorical variables
## 
#### Let us visualize the distribution of species within each island

```{r fig.width = 7.5, fig.asp = 0.4}
penguins |> 
  ggplot(aes(x=island, fill = species)) +
  geom_bar()

```

---
#### If you want to see species "distribution" across islands
#### Then do a relative frequency plot
##### Note: we map the variable that will be separated into bars to the x aesthetic, and the variable that will change the colors inside the bars to the fill aesthetic.

```{r fig.width = 7.5, fig.asp = 0.4}
penguins |> 
  ggplot(aes(x=island, fill = species)) +
  geom_bar(position = "fill")

```

---
## Two numeric variables
#### Scatterplot with geom_point() and smooth curves with geom_smooth()

```{r fig.width = 7.5, fig.asp = 0.5}

penguins |> 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

---
## Three or more variables

```{r fig.width = 7.5, fig.asp = 0.5}
penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = island))
```

---
## Facets
#### Too many aesthetic mappings in a plot makes it cluttered and confusing
#### Solution: Split plot into seperate plots, called facets
#### To facet a plot by a single variable, use facet_wrap()
#### The argument to facet_wrap is ~ followed by a categorical variable

---

```{r fig.width = 7.5, fig.asp = 0.5}

penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  facet_wrap(~island)

```

---

## Saving your plots

```{r fig.width = 7.5, fig.asp = 0.4}

penguins |> 
  ggplot(aes(flipper_length_mm,body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  facet_wrap(~island)
ggsave(filename = "testSaveOfPenguins.png")

```

