---
title: "Data Wrangling"
author: "Roy Thomas"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "Wilke-slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")

library(tidyverse)
library(palmerpenguins)
# ggplot2 settings so plots scale well for slide size 
theme_set(theme_gray(16))
update_geom_defaults("point", list(size = 2))
```

## Agenda

--

- ####Attendance
- ####Analytics in the News
- ####Review of past OYOTs
- ####Filter, select, arrange, count, mutate
- ####The pipe symbol
- ####Relocate, slice, join
- ####OYOTs (On Your Own Tasks)

---

##Review of past OYOTs

- ####R and RStudio is up to date?
- ####All packages installed?


---

## Elementary data manipulations

--

- Pick rows: `filter()`

--

- Pick columns: `select()`

--

- Sort rows: `arrange()`

--

- Count things: `count()`

--

- Make new columns: `mutate()`

---
class: center, middle

# But first: the pipe operator `%>%` or `|>`
(To enable the new operator, you may have to
go to Tools --> Global Options --> Code, and then
make sure that the checkbox for "Use native pipe 
operator, is checked.)

---
class: center, middle

# `%>%` and `|>` is pronounced "and then"
(The shortcut for the new pipe operator is 
Cntrl + Shift = M or Command + Shift + M)

---

## The pipe symbol feeds data into functions

.tiny-font[
```{r}
library(palmerpenguins) # loads the `penguins` dataset

head(penguins)
```
]

---

## The pipe `|>` feeds data into functions

.tiny-font[
```{r}
library(palmerpenguins) # loads the `penguins` dataset

# head(penguins)
penguins |>
  head()
```
]



---

## The pipe `|>` feeds data into functions

.tiny-font[
```{r penguins-scatter, eval = FALSE}
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color = species)) + geom_point()
```
]

---

## The pipe `|>` feeds data into functions

.tiny-font[
```{r penguins-scatter2, eval = FALSE}
# ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color = species)) + geom_point()

penguins |>
  ggplot(aes(bill_length_mm, bill_depth_mm, color = species)) + geom_point()
```
]


[//]: # "segment ends here"

---
class: center middle

## Picking rows or columns, and sorting

---

## Pick rows from a table: `filter()`

<br>

.center[
<img src = "dataWrangling_files/filter.svg", width = 75%></img>
]


---

## Filter out penguins of species Gentoo

.tiny-font[
```{r}
penguins |>
  filter(species == "Gentoo")
```
]

---

## Filter out penguins with bill length > 50 mm

.tiny-font[
```{r}
penguins |>
  filter(bill_length_mm > 50)
```
]

---

## Pick columns from a table: `select()`

<br>

.center[
<img src = "dataWrangling_files/select.svg", width = 75%></img>
]


---

## Pick columns `species`, `island`, and `sex`

.tiny-font[
```{r}
penguins |>
  select(species, island, sex)
```
]

---

## Sort the rows in a table: `arrange()`

<br>

.center[
<img src = "dataWrangling_files/arrange.svg", width = 75%></img>
]

---

## Sort penguins by bill length, ascending

.tiny-font[
```{r}
penguins |>
  arrange(bill_length_mm)
```
]

---

## Sort penguins by bill length, descending

.tiny-font[
```{r}
penguins |>
  arrange(desc(bill_length_mm))
```
]


[//]: # "segment ends here"

---
class: center middle

## Counting things


---

## Count things

.tiny-font[
```{r}
penguins
```
]

---

## Count things

.tiny-font[
```{r eval = FALSE}
penguins |>
  count(species)
```

```{r echo = FALSE}
x <- penguins |>
  count(species)
x
```
]

---

## Count things

.tiny-font[
```{r}
penguins |>
  count(species, island)
```
]

---

## Use the pipe to build analysis pipelines

.tiny-font[
```{r}
penguins |>
  filter(species == "Adelie")
```
]

---

## Use the pipe to build analysis pipelines

.tiny-font[
```{r}
penguins |>
  filter(species == "Adelie") |>
  select(island, sex)
```
]


---

## Use the pipe to build analysis pipelines

.tiny-font[
```{r}
penguins |>
  filter(species == "Adelie") |>
  select(island, sex) |>
  count(island, sex)
```
]


[//]: # "segment ends here"

---
class: center middle

## Adding new columns to a table

---
## Make a new table column: `mutate()`

<br>

.center[
<img src = "dataWranglingfiles/mutate.svg", width = 75%></img>
]

---

## Example: flipper length in cm

.tiny-font[
```{r}
penguins |>
  select(species, island, flipper_length_mm)
```
]

---

## Example: flipper length in cm

.tiny-font[
```{r}
penguins |>
  select(species, island, flipper_length_mm) |>
  mutate(flipper_length_cm = flipper_length_mm / 10)
```
]

---

## Make multiple columns at once

.tiny-font[
```{r}
penguins |>
  select(species, island, flipper_length_mm) |>
  mutate(
    flipper_length_cm = flipper_length_mm / 10,  # <- notice the comma
    flipper_length_in = flipper_length_mm / 25.4
  )
```
]


[//]: # "segment ends here"

---

## Further reading

- R for Data Science: [Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html)
- R for Data Science: [Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html)
- **dplyr** documentation: [Introduction to dplyr](https://dplyr.tidyverse.org/articles/dplyr.html)
- **dplyr** reference documentation: [One table verbs](https://dplyr.tidyverse.org/reference/index.html#section-one-table-verbs)

