---
title: "Data Wrangling2"
author: "Roy Thomas"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "Wilke-slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")

library(tidyverse)
# ggplot2 settings so plots scale well for slide size 
theme_set(theme_gray(16))
update_geom_defaults("point", list(size = 2))
```

## Agenda

--

- ####Attendance
- ####Analytics in the News
- ####Review of past OYOTs (data manipulation verbs)
- ####Data Exploration: Missing Values
- ####Data Exploration: Outliers
- ####OYOTs (On Your Own Tasks)

---

##Attendance

---

##Analytics in the News

---

##Review of past OYOTs

- ####Review of filter, select, arrange, count, mutate
- ####Review of relocate, slice

---

## Data Manipulation on the StarWars dataset
.tiny-font[
```{r}
library(tidyverse)  # This library has a LOT OF functions that we use
?starwars
starwars
View(starwars)

```
]

---

## Filter
.tiny-font[
```{r}
starwars |> 
  filter(skin_color == "light", eye_color == "brown")
```
]

---

#### starwars |> arrange(height, mass)
#### starwars |> arrange(desc(height))
#### starwars |> slice(5:10)
#### slice_head() and slice_tail()
#### slice_head(n=3)
#### slice_sample(n=5)

---

#### slice_min() and slice_max() select rows with highest or 
#### lowest values of a variable. Note that we first must choose 
#### only the values which are not NA.
#### starwars |> 
####   filter(!is.na(height)) |> 
####   slice_max(height, n = 3)
---

#### select allows you to zoom in to the columns of interest
#### starwars |> select(hair_color, skin_color, eye_color)
#### starwars |> select(hair_color:eye_color)
#### starwars |> select(!(hair_color:eye_color))
#### starwars |> select(ends_with("color"))

---

##### rename and mutate
##### starwars|> rename(home_world = homeworld)
##### starwars |> mutate(height_m = height / 100)
##### starwars |> 
#####   mutate(height_m = height / 100) %>%
#####   select(height_m, height, everything())
##### starwars |> 
#####   mutate(
#####     height_m = height / 100,
#####     BMI = mass / (height_m^2)
#####   ) |> 
#####   select(BMI, everything())
##### If you only want to keep the new variables, use .keep = "none"
##### starwars |> 
#####   mutate(
#####     height_m = height / 100,
#####     BMI = mass / (height_m^2),
#####     .keep = "none"
#####   )

---

##### Change column order with relocate
##### starwars|> 
#####   relocate(sex:homeworld, .before = height)
###Summarize
##### starwars|> 
#####   summarise(height = mean(height, na.rm = TRUE))

---

## Dealing with Missing Values
To identify missing values use is.na() which returns a logical vector with TRUE in the element locations that contain missing values represented by NA. is.na() will work on vectors, lists, matrices, and data frames.

##### vector with missing data
##### x <- c(1:4, NA, 6:7, NA)
##### x

##### is.na(x)

---
## data frame with missing data
##### df <- data.frame(col1 = c(1:3, NA),
#####                  col2 = c("this", NA,"is", "text"), 
#####                  col3 = c(TRUE, FALSE, TRUE, TRUE), 
#####                  col4 = c(2.5, 4.2, 3.2, NA),
#####                  stringsAsFactors = FALSE)

##### identify NAs in full data frame
##### is.na(df)

##### identify NAs in specific data frame column
##### is.na(df$col4)
---
To identify the location or the number of NAs we can leverage the which() and sum() functions:
##### identify location of NAs in vector
##### which(is.na(x))

##### identify count of NAs in data frame
##### sum(is.na(df))
## compute the total missing values in each column is to use colSums():
##### colSums(is.na(df))
---

## OYOTs (On Your Own Tasks)
#### https://thomasadventure.blog/posts/r-count-na/ (also shows heatmap of na’s)
#### Outliers: https://statsandr.com/blog/outliers-detection-in-r/

---

