---
title: "SCM 651 Visual Analytics Intro"
author: "Roy Thomas"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "Wilke-slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")
```

## Introduction

--

### What is Visual Analytics

--


.center[
<img src = "introduction_files/Visual-analytics-between-data-and-human.png", width = 55%, height = 55%></img>
]

.center[
https://www.researchgate.net/figure/Visual-analytics-as-the-interplay-between-data-analysis-visualization-and-human_fig2_333884128
]

---
## An internal dataset of R: anscombe
.xtiny-font[
```{r include=TRUE, echo=FALSE}
knitr::kable(anscombe)
```
]
---
.xtiny-font[
```{r include=TRUE, echo=TRUE, out.width='400px'}
library(ggplot2)
ggplot(data = anscombe, aes(x=x1,y=y1)) +
  geom_point()
```
]
---
.xtiny-font[
```{r include=TRUE, echo=TRUE, out.width='400px'}
library(ggplot2)
ggplot(data = anscombe, aes(x=x1,y=y1)) +
  geom_point()+geom_smooth(method = "lm")
```
]
---
.xtiny-font[
```{r include=TRUE, echo=TRUE, out.width='400px'}
library(ggplot2)
ggplot(data = anscombe, aes(x=x2,y=y2)) +
  geom_point()
```
]
---
.xtiny-font[
```{r include=TRUE, echo=TRUE, out.width='400px'}
library(ggplot2)
ggplot(data = anscombe, aes(x=x2,y=y2)) +
  geom_point()+geom_smooth(method = "lm")
```
]
---
.xtiny-font[
```{r include=TRUE, echo=TRUE, out.width='400px'}
library(ggplot2)
ggplot(data = anscombe, aes(x=x2,y=y2)) +
  geom_point()+geom_line()
```
]
---

## What is ggplot?

.center[
<img src = "introduction_files/ggplot2-website-title.png", width = 60%></img>
]

--

.center[
<img src = "introduction_files/ggplot2-website-footer.png", width = 60%></img>
]

[//]: # "segment ends here"

---

####Plot of mpg against disp - using the base function "plot" in R
```{R include=TRUE, echo=TRUE}
plot(mtcars$disp,mtcars$mpg)
```

---
class: middle, center

## Markdown

--

**Input:**

.small-font[
```
This is a sentence in Markdown, containing `code`, **bold text**,
and _italics_.
```
]


--

**Output:**

.small-font[
This is a sentence in Markdown, containing `code`, **bold text**, and _italics_.
]

---

##Formatting commands

The number of hashtags controls the level of the heading.<br>
You can have one to six hash tags

.pull-left[
**Input:**
```
#Heading1
##Heading2
###Heading3
####Heading4
#####Heading5
######Heading6
```
]

--
.pull-right[
**Output:**
#Heading1
##Heading2
###Heading3
####Heading4
#####Heading5
######Heading6
]
---

##Formatting commands

.pull-left[
**Input:**
```
 **bold**  
`code`  
 _italics_  
 **_bold and italics_**  
- unordered list
- item 2
 - item 2a (indent 1 tab)
 - item 2b
1. ordered list
2. item 2
 - item 2a (indent 1 tab)
 - item 2b
```
]

--
.pull-right[
**Output:**

**bold**  
`code`  
_italics_  
**_bold and italics_**  
- unordered list
- item 2
 - item 2a (indent 1 tab)
 - item 2b
1. ordered list
2. item 2
 - item 2a (indent 1 tab)
 - item 2b
]


---

## R Markdown combines Markdown and R code

--

**Input:**

.small-font[

````
The function `rnorm()` creates normal variates.
`r ''````{R}
rnorm(5) # create 5 normal variates
```
````
]

--

**Output:**

.small-font[

The function `rnorm()` creates normal variates.
```{R}
rnorm(5) # create 5 normal variates
```
]

[//]: # "segment ends here"

---

## Further reading

- [R Markdown tutorial](https://rmarkdown.rstudio.com/lesson-2.html)
- [R Markdown cheatsheet](https://github.com/rstudio/cheatsheets/raw/main/rmarkdown.pdf)
- [R Markdown book](https://bookdown.org/yihui/rmarkdown/)
