---
title: "Intro to RMarkdown"
author: "Roy Thomas"
date: "2023-08-28"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




---
#############################################################
#                                                           #
#   Click on "Knit" in RStudio to render this worksheet.    #
#                                                           #
#############################################################
title: "Intro to R Markdown"
author: "Roy Thomas"
output: html_document
---

R Markdown documents are documents that combine text, R code, and R output, including figures. They are a great way to produce self-contained and documented statistical analyses.

In this first worksheet, you will learn how to do some basic markdown editing. After you have made a change to the document, press "Knit HTML" in R Studio and see what kind of a result you get.

Edit only below this line.

---------------------------------------------

## 1. Basic Markdown editing

Try out basic R Markdown features, as described here: https://rmarkdown.rstudio.com/authoring_basics.html. 
- Write text that is in bold.
- Write text that is in italics.
- write text that is in different sized "headers"
- Create a bulleted list
- Create a numbered list
- Create a nested list

## 2. Embedding R code

R code embedded in R chunks will be executed and the output will be shown.



```{r}
# R code goes here
summary(mtcars)

x <- rnorm(100)  # random sample from normal distribution
dens <- density(x)    # calculate density


plot(dens)            # plot density
```

Now you try it. For example, take the built-in data set `cars`, which lists speed and stopping distance for cars from the 1920. Plot speed vs. distance. Then write a few sentences describing what you see.

Here are some hints:\
To figure out what the cars dataset is, use ?cars\
You can LOOK at the dataset with various commands, like "head", "glimpse", "summary"...  Remember that in each of these commands, you have to provide the name of the dataset as an argument.  

To do a simple plot, use the plot(x,y) command, where x and y are the names of the COLUMNS.  So, for instance, x would be cars$speed



