---
title: "Intro to RMarkdown"
author: "Roy Thomas"
date: "2023-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




---
#############################################################
#                                                           #
#   Click on "Knit" in RStudio to render this worksheet.    #
#                                                           #
#############################################################
title: "Intro to R Markdown"
author: "Roy Thomas"
output: html_document
---

R Markdown documents are documents that combine text, R code, and R output, including figures. They are a great way to produce self-contained and documented statistical analyses.

In this first worksheet, you will learn how to do some basic markdown editing. After you have made a change to the document, press "Knit HTML" in R Studio and see what kind of a result you get.

Edit only below this line.

---------------------------------------------

## 1. Basic Markdown editing

Try out basic R Markdown features, as described here: https://rmarkdown.rstudio.com/authoring_basics.html. 
- Write text that is in bold.
- Write text that is in italics.
- write text that is in different sized "headers"
- Create a bulleted list
- Create a numbered list
- Create a nested list

## 2. Embedding R code

R code embedded in R chunks will be executed and the output will be shown.



```{r}
# R code goes here
summary(mtcars)

x <- rnorm(100)  # random sample from normal distribution
dens <- density(x)    # calculate density


plot(dens)            # plot density
```

Now you try it. For example, take the built-in data set `cars`, which lists speed and stopping distance for cars from the 1920. Plot speed vs. distance. Then write a few sentences describing what you see.


